IT Spillovers in Long-Run TFP \\ A SVAR Approach

### 2018-01-21 Latex/Dissertation_Workshop_23Jan2018/dw23jan2018.tex
update of presentation

\title{The Slowdown of TFP \\ Exogenous and Endogenous Mechanisms}

@@

	\begin{itemize}
	\item The neoclassical tradition (RBC literature) in macro has treated total factor productivity (TFP) as exogenous.

	(King \& Rebelo 1999, Christiano, Eichenbaum \& Trabandt 2014, 2018)

@@

The IT sector is a separate sector that uses IT investment $S_{i,t}$ to expand $A_t$. Importantly, we assume that the IT sector is also augmented by $\Psi_t$, as well as by the sector-specific productivity $\lambda_t$.

\begin{equation}
V_{i,t} = \lambda_t \Psi_t f(S_{i,t}) 
\end{equation}

\hyperlink{it_sector}{\beamergotobutton{IT sector in detail}}

@@

The price level $P$ is an aggregate of the prices intermediate goods. These are determined as an FOC to 
\begin{equation*}
Y_t(s) = \Psi_t K_t(s)^{\alpha}L_t(s)^{1-\alpha}
\end{equation*}

\hspace{5cm} $\hookrightarrow P$ is a function of $\Psi_t$

@@

The price of IT good, $P^{IT}$, is determined as an FOC to

\begin{equation}
V_{i,t} = \lambda_t \Psi_t f(S_{i,t})
\end{equation}

\hspace{5cm}  $\hookrightarrow P^{IT}$ is a function of $\Psi_t$ \textbf{and} $\lambda_t$

@@

$\Rightarrow$ The two shocks imply different responses of relative prices $\frac{P^{IT}}{P}$!

\begin{itemize}
\item A news shock moves both $P$ and $P^{IT}$, leaving relative prices unchanged.
\item An IT productivity shock moves only $P^{IT}$, thus changing relative prices.
\end{itemize}

@@

\begin{equation}
\epsilon^{TFP}_t =   \underbrace{V_{t-k}}_\text{news shock} + \underbrace{IT_{t-k}}_\text{IT productivity shock}  + \underbrace{\varepsilon_t}_\text{surprise tech shock} 
\end{equation}

@@

    \item The IT shock maximizes the remaining FEV of future TFP;

@@

>>>>

\title{IT Spillovers in Long-Run TFP \\ A SVAR Approach}

@@

	\begin{itemize}
	\item The neoclassical tradition (RBC literature) in macro has treated total factor productivity (TFP) as exogenous, working against endogenous growth models

	(King \& Rebelo 1999, Christiano, Eichenbaum \& Trabandt 2014, 2018, vs. e.g. Romer 1990)

@@

IT sector is a separate sector that uses IT investment and its own sectoral productivity $\lambda_t$ to expand $A_t$.  \hyperlink{it_sector}{\beamergotobutton{IT sector in detail}}

$\rightarrow$ IT sector is driven by profits from selling the IT good, but its advances spill over to aggregate TFP.

@@

Let $P$ be the price level (price of final good), $P^{IT}$ the price of the IT good and relative prices be $P^{IT}/P$.

@@

Intuition:
\begin{itemize}
\item News shock makes all sectors of the economy more productive $\rightarrow$ impacts the sectoral price $P^IT$ as well as P $\Rightarrow$ relative prices do not move.
\item IT productivity shock makes only the IT sector more productive $\rightarrow$ impacts the sectoral price $P^{IT}$ \textbf{only} $\Rightarrow$ relative prices move! 
\end{itemize}

@@

$\Rightarrow$ identifying restriction: news shock does not move relative prices.
\newline Accounting for price rigidities: news shock does not move relative prices after price adjustment has taken place (6-12 quarters).

Similar identification scheme: Fisher 2006.

%The price level $P$ is an aggregate of the prices intermediate goods. These are determined as an FOC to 
%\begin{equation*}
%Y_t(s) = \Psi_t K_t(s)^{\alpha}L_t(s)^{1-\alpha}
%\end{equation*}
%
%\hspace{5cm} $\hookrightarrow P$ is a function of $\Psi_t$ 
%
%\
%
%The price of IT good, $P^{IT}$, is determined as an FOC to
%
%\begin{equation}
%V_{i,t} = \lambda_t \Psi_t f(S_{i,t})
%\end{equation}
%
%
%\hspace{5cm}  $\hookrightarrow P^{IT}$ is a function of $\Psi_t$ \textbf{and} $\lambda_t$	
%
%\
%
%$\Rightarrow$ The two shocks imply different responses of relative prices $\frac{P^{IT}}{P}$!
%
%\begin{itemize}
%\item A news shock moves both $P$ and $P^{IT}$, leaving relative prices unchanged.
%\item An IT productivity shock moves only $P^{IT}$, thus changing relative prices.
%\end{itemize}

@@

Innovation in TFP at time $t$ can be decomposed as:

\begin{equation}
\epsilon^{TFP}_t =   \underbrace{V_{t-k}}_\text{news shock} + \underbrace{f(IT_{t-j})}_\text{IT productivity shock}  + \underbrace{\varepsilon_t}_\text{surprise tech shock} 
\end{equation}

@@

    \item The IT productivity shock maximizes the remaining FEV of future TFP;

@@

\item Show the identification assumption in the structural model.
